A short note on the penalty flux parameter for first order discontinuous Galerkin formulations

### 2017-05-03 docs/dg_penalty/test.tex
updating docs for penalty note + elasticity

%\documentclass{amsart}
\documentclass[preprint,10pt]{elsarticle}
%\usepackage{lineno}

%\documentclass[10pt]{article}
%\documentclass{siamart0216}

\usepackage{fullpage}
\usepackage{hyperref}
%\let\proof\relax
%\let\endproof\relax
\usepackage{amsmath,amssymb,amsfonts,mathrsfs}%,amsthm}
\usepackage[titletoc,toc,title]{appendix}

\usepackage{array} 
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage[textsize=footnotesize,color=green]{todonotes}
\usepackage{bm}
%\usepackage{tikz}
\usepackage[normalem]{ulem}
\usepackage{hhline}

%% ====================================== alg package
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algorithmicx}
\algblock{ParFor}{EndParFor}
% customising the new block
\algnewcommand\algorithmicparfor{\textbf{parfor}}
\algnewcommand\algorithmicpardo{\textbf{do}}
\algnewcommand\algorithmicendparfor{\textbf{end\ parfor}}
\algrenewtext{ParFor}[1]{\algorithmicparfor\ #1\ \algorithmicpardo}
\algrenewtext{EndParFor}{\algorithmicendparfor}
%% ====================================== end alg package

\usepackage{graphicx}
\usepackage{subfig}
\usepackage{color}

%% ====================================== graphics

\usepackage{pgfplots}
\usepackage{pgfplotstable}
\definecolor{markercolor}{RGB}{124.9, 255, 160.65}
\pgfplotsset{width=10cm,compat=1.9}
\pgfplotsset{
tick label style={font=\small},
label style={font=\small},
legend style={font=\small}
}

\usetikzlibrary{calc}
\usepackage{stmaryrd}

\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

\newcommand{\vect}[1]{\ensuremath\boldsymbol{#1}}
\newcommand{\tensor}[1]{\underline{\bm{#1}}}
\newcommand{\del}{\triangle}
\newcommand{\curl}{\grad \times}
\renewcommand{\div}{\grad \cdot}
\newcommand{\td}[2]{\frac{{\rm d}#1}{{\rm d}{\rm #2}}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\pdd}[2]{\frac{\partial^2#1}{\partial#2^2}}

\newcommand{\bs}[1]{\boldsymbol{#1}}

\newcommand{\equaldef}{\stackrel{\mathrm{def}}{=}}

\newcommand{\tablab}[1]{\label{tab:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}

\newcommand{\theolab}[1]{\label{theo:#1}}
\newcommand{\theoref}[1]{\ref{theo:#1}}
\newcommand{\eqnlab}[1]{\label{eq:#1}}
\newcommand{\eqnref}[1]{\eqref{eq:#1}}
\newcommand{\seclab}[1]{\label{sec:#1}}
\newcommand{\secref}[1]{\ref{sec:#1}}
\newcommand{\lemlab}[1]{\label{lem:#1}}
\newcommand{\lemref}[1]{\ref{lem:#1}}

\newcommand{\mi}{\mathrm{i}} %% roman "i"
\newcommand{\mb}[1]{\mathbf{#1}}
\newcommand{\mbb}[1]{\mathbb{#1}}
\newcommand{\mc}[1]{\mathcal{#1}}
\newcommand{\nor}[1]{\left\| #1 \right\|}
\newcommand{\snor}[1]{\left| #1 \right|}
\newcommand{\LRp}[1]{\left( #1 \right)}
\newcommand{\LRs}[1]{\left[ #1 \right]}
\newcommand{\LRa}[1]{\left\langle #1 \right\rangle}
\newcommand{\LRb}[1]{\left| #1 \right|}
\newcommand{\LRc}[1]{\left\{ #1 \right\}}
\newcommand{\LRceil}[1]{\left\lceil #1 \right\rceil}

\newcommand{\Grad} {\ensuremath{\nabla}}
\newcommand{\Div} {\ensuremath{\nabla\cdot}}
\newcommand{\Nel} {\ensuremath{{N^\text{el}}}}
\newcommand{\jump}[1] {\ensuremath{\llbracket#1\rrbracket}}
\newcommand{\avg}[1] {\ensuremath{\LRc{\!\{#1\}\!}}}
\newcommand{\uh}{\widehat{u}}
\newcommand{\Bh}{\widehat{B}}
\newcommand{\fnh}{\widehat{f}_n}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\Lk}{L^2\LRp{D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}
\newcommand{\pO}{\partial\Omega}
\newcommand{\Gh}{\Gamma_h}
\newcommand{\Gm}{\Gamma_{-}}
\newcommand{\Gp}{\Gamma_{+}}
\newcommand{\Go}{\Gamma_0}
\newcommand{\Oh}{\Omega_h}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{proposition}[theorem]{Proposition}
%\newtheorem{corollary}[theorem]{Corollary}

%\newenvironment{definition}[1][Definition]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
%\newenvironment{example}[1][Example]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\newcommand{\eval}[2][\right]{\relax
  \ifx#1\right\relax \left.\fi#2#1\rvert}

\def\etal{{\it et al.~}}

\def\arr#1#2#3#4{\left[
\begin{array}{cc}
#1 & #2\\
#3 & #4\\
\end{array}
\right]}
\def\vectwo#1#2{\left[
\begin{array}{c}
#1\\
#2\\
\end{array}
\right]}
\def\vecthree#1#2#3{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
\end{array}
\right]}
\def\vectfour#1#2#3#4{\left[
\begin{array}{c}
#1\\
#2\\
#3\\
#4\\
\end{array}
\right]}

\newcommand{\G} {\Gamma}
\newcommand{\Gin} {\Gamma_{in}}
\newcommand{\Gout} {\Gamma_{out}}

\newcommand{\note}[1]{{\color{blue}#1}}
\newcommand{\remark}[1]{\textbf{\color{red}#1}}

\newcommand{\tri}{{\rm tri}}
\newcommand{\sqr}{{\rm quad}}

\newcommand{\refhex}{\widehat{\mathcal{H}}}
\newcommand{\reftet}{\widehat{\mathcal{T}}}
\newcommand{\refwedg}{\widehat{\mathcal{W}}}

\newcommand{\refpyr}{\widehat{\mathcal{P}}}
\newcommand{\refpyrf}{\widehat{\mathcal{P}}^f}
\newcommand{\refpyrfq}{\widehat{\mathcal{P}}^{\sqr}}
\newcommand{\refpyrft}{\widehat{\mathcal{P}}^{\tri}}

\newcommand{\hex}{{\mathcal{H}}}
\newcommand{\tet}{{\mathcal{T}}}
\newcommand{\wedg}{{\mathcal{W}}}

\newcommand{\pyr}{{\mathcal{P}}}
\newcommand{\pyrf}{{\mathcal{P}}^f}
\newcommand{\pyrfq}{{\mathcal{P}}^{\sqr}}
\newcommand{\pyrft}{{\mathcal{P}}^{\tri}}

\newcommand{\LinfDk}{L^{\infty}\LRp{D^k}}

\newcommand{\diag}[1]{{\rm diag}\LRp{#1}}

\newcommand{\half}{1/2}

\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

%% d in integrand
\newcommand*\diff[1]{\mathop{}\!{\mathrm{d}#1}}

\makeatletter
\renewcommand\d[1]{\mspace{6mu}\mathrm{d}#1\@ifnextchar\d{\mspace{-3mu}}{}}
\makeatother

\newcommand{\squareface}{[-1,1]^2}

\date{}
%\author{Jesse Chan, T. Warburton}
%\title{A short note on the penalty flux parameter for first order discontinuous Galerkin formulations}

@@

%\maketitle
\begin{frontmatter}
\title{A short note on the penalty flux parameter for \note{discontinuous Galerkin formulations of first order hyperbolic systems}}

\author[rice]{Jesse Chan\corref{cor1}}
\ead{Jesse.Chan@caam.rice.edu}
\cortext[cor1]{Principal Corresponding author}
\author[vt]{T. Warburton}
\ead{tcew@vt.edu}
\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
\address[vt]{Department of Mathematics, Virginia Tech, 225 Stanger Street, Blacksburg, VA 24061-0123}

\address{}

\begin{abstract}
Penalty fluxes are dissipative numerical fluxes for high order discontinuous Galerkin (DG) methods which depend on a penalization parameter \cite{warburton2013low, ye2016discontinuous}.  We investigate the dependence of the spectra of high order DG discretizations on this parameter, and show that as its value increases, the \note{spectrum} of the DG discretization splits into two disjoint sets of eigenvalues.  One set converges to the eigenvalues of a conforming discretization, while the other set corresponds to spurious eigenvalues which are damped proportionally to $\tau$.  Numerical experiments also demonstrate that undamped spurious modes present in both in the limit of zero and large penalization parameters are damped for moderate values of the upwind parameter.  
\end{abstract}
\end{frontmatter}

\section{Introduction}

High order discontinuous Galerkin (DG) methods are a popular choice of discretization for simulations of time-domain wave propagation, due to their low numerical dispersion and their ability to accomodate unstructured meshes \cite{grote2006discontinuous,hesthaven2007nodal}.  For PDEs in second order form, a common choice of DG method is the symmetric interior penalty (SIPG) discontinuous Galerkin method \cite{riviere2008discontinuous}, which involves a penalization parameter $\tau$.  The role of this penalty parameter was investigated in \cite{Warburton20063205} for discretizations of the Maxwells equations.  There, it was shown that in the limit of $\tau\rightarrow\infty$, the \note{spectrum} of the DG discretization matrix splits into two disjoint sets of eigenvalues, with one set converging to eigenvalues of a conforming discretization and one set associated with divergent spurious eigenvalues.

We consider penalty fluxes for discretizations of first order systems of hyperbolic PDEs, which depend on some parameter $\tau \geq 0$.  For $\tau = 0$, these fluxes reduce to the energy-conserving and non-dissipative central fluxes \cite{fezoui2005convergence}.  For $\tau > 0$, penalty fluxes are related to upwind fluxes for problems with continuous coefficients.  We extend the results of \cite{Warburton20063205} to DG formulations for first order systems with penalty fluxes and show that the \note{spectrum} of the DG discretization matrix again splits into two disjoint sets of eigenvalues as $\tau\rightarrow\infty$.  The first set of eigenvalues correspond to the eigenvalues of a conforming discretization, while the second set of eigenvalues have real parts approaching $-\infty$ as $\tau\rightarrow \infty$ and correspond to damped spurious modes \note{which are non-conforming}.  Numerical experiments verify these results for the advection and acoustic wave equations in one and two dimensions.  Numerical experiments also show that, while increasing $\tau$ pushes most eigenvalues further left of the imaginary axis, the real parts of certain eigenvalues converge to zero as $\tau\rightarrow\infty$.  This implies that taking a moderate penalty parameter $\tau = 1$ damps spurious \note{\emph{conforming}} modes which remain otherwise undamped in time-dependent simulations with conforming discretizations.  \note{We show that penalized DG methods address both \textit{non-conforming} spurious modes present when using central fluxes ($\tau = 0$) and \textit{conforming} spurious modes which are present under conforming discretizations (or equivalently, under large penalty parameters $\tau \rightarrow \infty$).}

\section{Semi-discrete discontinuous Galerkin formulation}

We consider linear first order hyperbolic systems of PDEs in $\mathbb{R}^d$
\begin{equation}
\pd{\bm{U}}{t} + \sum_{i=1}^d \bm{A}_i\pd{{\bm{U}}}{\bm{x}_i} = 0
\label{eq:cons}
\end{equation}
with solution components $\bm{U} = (u_1, \ldots, u_M)^T$ and matrices $\bm{A}_i \in \mathbb{R}^{M\times M}$.  
\note{
For hyperbolic systems, any linear combination of $\bm{A}_i$ 
\begin{equation}
\bm{A}_{\alpha} = \sum_{i=1}^d \bm{\alpha}_i \bm{A}_i
\label{eq:hyperbolic}
\end{equation}
is diagonalizable with real eigenvalues.  
}

We assume the domain $\Omega$ is triangulated exactly using a mesh $\Oh$ consisting of elements $D^k$.  We define the $L^2$ inner product over elements $D^k$ and their boundaries $\partial D^k$ as
\[
\LRp{u,v}_{D^k} = \int_{D^k} u(\bm{x})v(\bm{x}) \diff x, \qquad \LRa{u,v}_{\partial D^k} = \int_{\partial D^k} u(\bm{x})v(\bm{x}) \diff x.
\]
A semi-discrete DG formulation for (\ref{eq:cons}) may be written as 
\begin{equation}
\sum_{D^k \in \Oh} \LRp{\pd{\bm{U}}{t},\bm{V} }_{\Lk} = \sum_{D^k \in \Oh}\LRp{ \note{\LRp{\sum_{i=1}^d\bm{A}_i \bm{U}, \pd{\bm{V}}{\bm{x}_i}}_{\Lk}} - \LRa{(\bm{A}_n\bm{U})^*,\bm{V}}_{\partial D^k}}
\label{eq:formulation}
\end{equation}
where $(\bm{A}_n\bm{U})^*$ is a numerical flux defined on boundary faces and shared faces between elements.

\note{
In the main proof of this paper, we make the assumption that the matrices $\bm{A}_i$ are symmetric and spatially constant for clarity and simplicity of presentation.  However, the numerical fluxes used are more generally applicable (for example, when $\bm{A}_i$ is non-symmetric, spatially varying, or discontinuous), and we present generalizations beyond symmetric and constant $\bm{A}_i$ where possible.  
%More specifically, we 
%\\
%Specify what assumptions are made on $\bm{A}_i$ and why.
%\begin{enumerate}
%\item Symmetric - 
%\item Locally constant - energy stable.
%\item Spatially constant - for simplicity and clarity.
%\end{enumerate}
%For the remainder of this paper, we will also assume that $\bm{A}_i$ are piecewise constant.  
}

Finally, we note that
%, while the assumptions on constant $\bm{A}_i$ are restrictive, 
the results in this work are straightforward to generalize to the following modification of (\ref{eq:cons})
\begin{equation}
\bm{A}_0\pd{\bm{U}}{t} + \sum_{i=1}^d \bm{A}_i\pd{{\bm{U}}}{\bm{x}_i} = \bm{f}.
\label{eq:consA}
\end{equation}
We assume $\bm{A}_0 = \bm{I}$ in this work for brevity; however, all results are straightforward to extend to systems of the form for any $\bm{A}_0 = \bm{A}_0(\bm{x})$ which may vary spatially and is pointwise symmetric and positive-definite.  Many models of wave propagation in general heterogeneous media can be represented in the form of (\ref{eq:consA}), including the acoustic wave equation \cite{chan2016weight1}, linear elasticity \cite{de2008interior,wilcox2010high,ye2016discontinuous}, and Maxwell's equations \cite{hesthaven2002nodal,hesthaven2004high,grote2007interior,warburton2013low}.  %Results also follow for matrices $\bm{A}_i$ may be symmetrized through an appropriate change of variables \cite{kopriva2014energy}.  
\note{Additionally, we note that many hyperbolic systems of interest are symmetrizable, and can be cast into the form of (\ref{eq:consA}) with symmetric $\bm{A}_i$ through an appropriate change of variables \cite{kopriva2014energy}.  }

\subsection{Penalization terms in numerical fluxes}
\label{sec:scaletau}

Let $f$ be a face \note{shared between} an element $D^{k,-}$ and its neighbor $D^{k,+}$.  We denote the outward normal on the face $f$ of $D^{k,-}$ as $\bm{n}$, and let $\bm{U}^+, \bm{U}^-$ be evaluations of $\bm{U}$ restricted to $D^{k,+}$ and $D^{k,-}$, respectively.  The jump and average scalar-valued functions are defined as 
\[
\note{\jump{u} = u^+ - u^-, \qquad \avg{u} = \frac{{u}^+ + u^-}{2}.}
\]
We define the normal matrix across a face as
\[
{\bm{A}}_n = \sum_{i=1}^d {\bm{A}_i\bm{n}_i}.
\]
\note{We first introduce the penalty flux for continuous coefficients (where the value of $\bm{A}_n$ is single valued across shared faces) as}
\[
(\bm{A}_n\bm{U})^* = \bm{A}_n\avg{\bm{U}} - \frac{\tau}{2} \bm{A}_n^T \bm{A}_n\jump{\bm{U}}, \qquad \tau \geq 0.
\]
We compare the penalty flux to the well-known upwind numerical flux.  For \note{continuous coefficients, the upwind flux can be written using an eigenvalue decomposition of $\bm{A}_n$ (see, for example, \cite{ye2016discontinuous} and Chapter 2.4 of \cite{hesthaven2007nodal})}.  
%By symmetry of $\bm{A}_i$, ${\bm{A}}_n$ contains real eigenvalues, and admits an eigenvalue decomposition
Under assumption (\ref{eq:hyperbolic}), ${\bm{A}}_n$ has real eigenvalues, and admits an eigenvalue decomposition
\[
\bm{A}_n = \bm{V}{\bm{\Lambda}}\note{\bm{V}^{-1}}, \qquad \bm{\Lambda} = 
\left(\begin{array}{ccc}
\lambda_1 & & \\
& \ddots & \\
& & \lambda_d
\end{array}\right)
\]
For problems with continuous coefficients, the upwind numerical flux over a face $f \in \Gh$ can then be defined  as
\[
(\bm{A}_n\bm{U})^* = \bm{A}_n^+\bm{U}^- + \bm{A_n}^- \bm{U}^+,
\]
where the matrices $\bm{A_n}^+,\bm{A_n}^-$ are constructed from the positive and negative eigenvalues 
\begin{align*}
\bm{A}_n^+ &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} + \LRb{\bm{\Lambda}}} \note{\bm{V}^{-1}}\\
\bm{A}_n^- &= \frac{1}{2}\bm{V} \LRp{\bm{\Lambda} - \LRb{\bm{\Lambda}}} \note{\bm{V}^{-1}},
\end{align*}
and $\LRb{\bm{\Lambda}}$ is the diagonal matrix whose entries consist of the absolute values of the eigenvalues $\LRb{\lambda_i}$.  This can be rewritten as the sum of the central flux and a penalization
\[
(\bm{A}_n\bm{U})^* = \bm{A}_n\avg{\bm{U}} - \frac{1}{2}\bm{V}\LRb{\bm{\Lambda}}\note{\bm{V}^{-1}} \jump{\bm{U}}.  
\]
\note{
Assuming constant, symmetric $\bm{A}_i$ and a penalty numerical flux, one can show the semi-discrete energy estimate \cite{hesthaven2007nodal, warburton2013low, kopriva2014energy, chan2016weight1} for (\ref{eq:cons})
\[
\sum_{D^k\in \Oh}\pd{}{t}\nor{\bm{U}}^2_{L^2\LRp{D^k}} = -\sum_{D^k \in \Oh} \frac{\tau}{2}\nor{\bm{A}_n\jump{\bm{U}}}^2_{L^2\LRp{\partial D^k}} \leq 0.
\]
The penalization term can thus be interpreted as adding a stabilizing contribution which dissipates components of $\bm{U}$ for which $\bm{A}_n\jump{\bm{U}}$ is large, with larger values of $\tau$ increasing the rate of dissipation (this is true even when $\bm{A}_n$ is not symmetric).  We will relate this dissipation to the spectrum of the DG discretization matrix in Section~\ref{sec:spec}.  In particular, we show that the real parts of eigenvalues corresponding to components of eigenvectors with $\bm{A}_n\jump{\bm{U}}\neq 0$ approach $-\infty$ as $\tau \rightarrow \infty$.

When $\bm{A}_n$ is not continuous across faces, the penalty numerical flux can be generalized as follows \cite{ye2016discontinuous}
\[
(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \frac{\tau}{2} \bm{A}_n^T \jump{\bm{A}_n\bm{U}},
\]
where we define the normal jump and average
\[
\jump{\bm{A}_n\bm{U}}\coloneqq \sum_{i=1}^d\bm{n}_i \jump{ \bm{A}_i \bm{U}}, \qquad \avg{\bm{A}_n\bm{U}} \coloneqq \sum_{i=1}^d\bm{n}_i\avg{ \bm{A}_i \bm{U}}.
\]
If $\bm{A}_i$ is constant over each element $D^k$, we again have an energy estimate with a dissipative contribution from the penalty flux
\[
\sum_{D^k\in \Oh}\pd{}{t}\nor{\bm{U}}^2_{L^2\LRp{D^k}} = -\sum_{D^k \in \Oh} \frac{\tau}{2}\nor{\jump{\bm{A}_n\bm{U}}}^2_{L^2\LRp{\partial D^k}} \leq 0.
\]
The upwind flux can similarly be extended for discontinuous coefficients
\[
(\bm{A}_n\bm{U})^* = \avg{\bm{A}_n\bm{U}} - \frac{1}{2} \jump{\bm{V}\LRb{\Lambda}\bm{V}^{-1}\bm{U}},
\]
where the eigenvalue decompositions inside the jump is evaluated as 
\begin{align*}
&\jump{\bm{V}\LRb{\Lambda}\bm{V}^{-1}\bm{U}} = {\bm{V}^+\LRb{\Lambda}^+\LRp{\bm{V}^+}^{-1}\bm{U}^+} - {\bm{V}^-\LRb{\Lambda}^-\LRp{\bm{V}^-}^{-1}\bm{U}^-}\\
&\bm{A}_n^+ = \sum_{i=1}^d \bm{n}_i \bm{A}_i^+ = \bm{V}^+{\Lambda}^+\LRp{\bm{V}^+}^{-1}, \qquad \bm{A}_n^- = \sum_{i=1}^d \bm{n}_i \bm{A}_i^- = \bm{V}^-{\Lambda}^-\LRp{\bm{V}^-}^{-1}.
\end{align*} 
Finally, we note that the energy estimate changes when $\bm{A}_i$ varies over an element, though the penalty flux still adds a dissipative contribution to the right hand side \cite{kopriva2014energy}.  
}

\subsection{Scaling of $\tau$}

We note that including a positive-definite weight matrix $\bm{W}$ into the penalty term
\[
\frac{\tau}{2} \bm{A}_n^T \bm{W}\bm{A}_n\jump{\bm{U}}
\]
recovers the upwind fluxes for $\tau = 1$ and $\bm{W} = \bm{V}\LRb{\bm{\Lambda}}^{\dagger}\bm{V}^{-1}$, where $\bm{A}^{\dagger}$ is the pseudo-inverse of $\bm{A}$ \cite{horn2012matrix}.  However, the advantage of penalty fluxes is that, unlike upwind fluxes, they can be used when explicit knowledge of a \note{joint} diagonalization of $\bm{A}_n$ is not known, such as for linear elasticity in anisotropic media \cite{ye2016discontinuous}.  Thus, we assume $\bm{W} = \bm{I}$ in this work.  
% instead seek an appropriate scaling $\tau$ such that $\bm{V}\LRb{\bm{\Lambda}}\note{\bm{V}^{-1}}$ and $\tau\bm{A}_n^T\bm{A}_n$ are of the same magnitude.  
%For $\bm{A}_i$ symmetric, this implies that $\tau$ should be proportional to $1/\LRp{\max_i\LRb{\lambda_i}}$.  
%This ensures that the penalty term is of the same magnitude as the upwind penalization term, and that CFL restrictions in explicit time-stepping methods will be \note{similar} for both upwind and penalty fluxes.

\note{It remains to specify an appropriate value of $\tau$.  We first note that the timestep is proportional to the spectral radius of the DG matrix, which increases proportionally to $\tau$ \cite{chan2015gpu}.  Since both upwind and penalty fluxes are identical apart from the penalization term, we choose $\tau$ such that the penalty flux and upwind penalization terms are of the same magnitude 
\[
\nor{\frac{\tau}{2} \bm{A}_n^T\bm{A}_n} = \nor{\frac{1}{2}\bm{V}\LRb{\bm{\Lambda}}{\bm{V}^{-1}}}.
\]
This is sufficient to ensure that the spectral radius of the DG matrix is similar when using either upwind or penalty fluxes.

The penalty flux and upwind penalization terms can both be bounded from above in terms of the largest magnitude eigenvalue $\max_i\LRb{\lambda_i}$ and the condition number $\kappa\LRp{\bm{V}}$ of the eigenvector matrix $\bm{V}$
\[
\nor{\frac{1}{2}\bm{V}\LRb{\bm{\Lambda}}{\bm{V}^{-1}}} \leq \max_i \LRb{\lambda_i} \kappa\LRp{\bm{V}}, \qquad \nor{\bm{A}_n^T\bm{A}_n} = \nor{\bm{V}^{-T}\bm{\Lambda}\bm{V}^T \bm{V}\bm{\Lambda}\bm{V}^{-1}} \leq \max_i\LRb{\lambda_i}^2\kappa\LRp{\bm{V}}^2.
\]
Thus, choosing $\tau = 1/\LRp{\max_i\LRb{\lambda_i}\kappa\LRp{\bm{V}}}$ results in a penalization term whose magnitude is the same as that of the upwind penalization term.  When $\bm{A}_n$ is symmetric (or symmetrizable through a change of variables), $\bm{V}$ is unitary and $\kappa\LRp{\bm{V}} = 1$, simplifying the selection of $\tau$ to $\tau = 1/\max_i\LRb{\lambda_i}$.  
%Since $\bm{A}_n$ is symmetric, $\bm{V}$ is unitary and $\nor{\bm{V}\LRb{\bm{\Lambda}}\bm{V}^{-1}} = \nor{{\bm{\Lambda}}}$.  Thus, the magnitude of the upwind penalization term is $\max_i\LRb{\lambda_i}$.  For penalty fluxes, the penalization term is 
%\[
%\nor{\bm{A}_n^T\bm{A}_n} = \nor{\bm{V}{\bm{\Lambda}}^2\bm{V}^{-1}} = \nor{{\bm{\Lambda}}^2} = \max_{i} \lambda_i^2,
%\]
%where we have again used symmetry of $\bm{A}_n$.  Thus, choosing $\tau = 1/\LRp{\max_i\LRb{\lambda_i}}$ results in a penalization term whose magnitude is the same as that of the upwind penalization term.  For non-symmetric $\bm{A}_n$, an appropriate scaling of $\tau$ can be determined based on the singular values of $\bm{A}_n$.  
}

\section{Dependence of \note{spectrum} on the penalty parameter}
\label{sec:spec}
The remainder of this work focuses on the influence of $\tau$ on the \note{spectrum} of DG discretization matrices.  We will show that, as $\tau$ increases, the \note{spectrum} splits into two sets of eigenvalues corresponding respectively to conforming and spurious modes, where conformity is defined based on $\bm{A}_n$.  \note{For clarity and simplicity of presentation, we will assume that $\bm{A}_i$ are symmetric and spatially constant for the remainder of this paper.  }%We will also assume that all elements $D^k$ are affine mappings of a reference element $\widehat{D}$, which will ensure that }

\subsection{Conforming and non-conforming approximation spaces}

Let $V$ denote the piecewise polynomial approximation space \note{over an element $D^k$}
\[
V = \LRc{ \bm{U} \in\LRp{ L^2(\Omega)}^d : \left.\bm{U}\right|_{D^k} \in \LRp{P^N(D^k)}^d, \quad \note{\forall D^k \in \Oh}}.  
\]
We decompose $V$ into a conforming space $V^C$ and a non-conforming space $V^{NC}$ based on the penalization term.  The conforming space is defined to be elements of $V$ for which the penalization term vanishes over all faces of all elements.  For penalty fluxes, \note{this term is $\bm{A}_n^T \bm{A}_n\jump{\bm{U}}$, such that} $V^C$ is
\begin{equation}
V^{C} = \LRc{ \bm{U}(\bm{x}) \in V: \quad  \bm{A}_n\jump{\bm{U}} = 0, \quad \forall f \in \partial D^k, \quad \forall D^k\in \Oh}.
\label{eq:conf}
\end{equation}
\note{This simplification can be made due to the fact that the null space of $\bm{A}_n^T\bm{A}_n$ is identical to the null space of $\bm{A}_n$.\footnote{\note{The null space of $\bm{A}_n^T\bm{A}_n$ contains the null space of $\bm{A}_n$.  However, if $\bm{A}_n^T\bm{A}_n\bm{U} = 0$, then $\bm{U}^T\bm{A}_n^T\bm{A}_n\bm{U} = \nor{\bm{A}_n\bm{U}}^2 = 0$ as well, implying that the null space of $\bm{A}_n$ also contains the null space of $\bm{A}_n^T\bm{A}_n$.
}}
}  
The non-conforming approximation space $V^{NC}$ is defined as $L^2$ orthogonal complement of $V^{C}$ in $V$
\[
V^{NC} = \LRc{ \bm{U}(\bm{x}) \in V: \LRp{\bm{u},\bm{v}}_{L^2(\Omega)} = 0, \quad \forall \bm{v}\in V^C}.
\]
A conforming space could also be defined using the upwind flux by seeking $\bm{U}$ such that $\bm{V}\LRb{\bm{\Lambda}}\note{\bm{V}^{-1}} \jump{\bm{U}} = 0$.  When the coefficient matrices $\bm{A}_i$ are spatially continuous, the condition $\bm{V}\LRb{\bm{\Lambda}}\note{\bm{V}^{-1}} \jump{\bm{U}} = 0$ and the penalty flux condition $\bm{A}_n^T \bm{A}_n\jump{\bm{U}} = 0$ are both equivalent to
\[
\bm{v}_i^T \jump{\bm{U}} = 0, \qquad \lambda_i \neq 0,
\]
where $\lambda_i, \bm{v}_i$ are an eigenvalue and eigenvector of $\bm{A}_n$.  As a result, the conforming and non-conforming spaces induced by upwind and penalty fluxes are identical.

We note that, for problems with discontinuous material data, it is less straightforward to show equivalence between the conforming spaces $V^C$ resulting from upwind and penalty fluxes.  However, penalty fluxes can still be applied in the presence of discontinuous coefficients by incorporating these spatial variations into the matrix $\bm{A_0}$.  When paired with penalty fluxes, this approach results in energy stable DG methods \cite{chan2016weight1}.  Additionally, DG methods with penalty fluxes are observed to perform similarly to DG methods with upwind fluxes for several realistic wave propagation problems with discontinuous coefficients \cite{ye2016discontinuous}.

\subsection{Scalar advection and acoustic wave propagation}
\label{sec:confexamples}
We illustrate the conforming spaces induced by penalty fluxes using the scalar advection equation and acoustic wave equation as examples.  The scalar advection equation is given as
\[
\pd{u}{t} + \Div\LRp{\bm{\beta}u} = 0.
\]
where $\bm{\beta}$ is the direction of advection.  \note{We also assume $\bm{\beta}$ is constant, such that the resulting equation can be rearranged into the form (\ref{eq:cons}).}  The penalty flux is 
\[
(\bm{A}_n\bm{U})^* = \bm{\beta}_n\avg{ u} - \tau\frac{\LRp{\bm{\beta}_n}^2}{2}\jump{u}.
\]
\note{Setting $\tau = \LRb{\bm{\beta}_n}$ recovers the standard upwind flux.}  
The conforming space $V^C$ induced by this flux is then
\[
V^C = \LRc{ u \in L^2\LRp{\Omega} : \quad \left.u\right|_{D^k} \in P^N(D^k), \quad \LRb{\bm{\beta}_n}\jump{u} = 0}.
\]
In one dimension, this simply implies that $u$ is $C^0$-continuous across element boundaries.  In higher dimensions, this implies that $u$ is continuous along streamlines or directions $\bm{d}(\bm{x})$ where $\bm{\beta}\cdot \bm{d} \neq 0$.

>>>>

%\documentclass[preprint,10pt]{elsarticle}
\documentclass[10pt]{article}
%\documentclass[review]{siamart0216}

\date{}

@@

\maketitle

One set converges to the eigenvalues of a conforming discretization, while the other set corresponds to spurious eigenvalues which are damped proportionally to $\tau$.  Numerical experiments also demonstrate that undamped spurious modes present in both in the limit of zero and large penalization parameters are damped for moderate values of the upwind parameter.  
\end{document}
